---
title: "Capstone-presentation"
author: "HarrySecor"
date: "11/30/2022"
output: html_document
---

## Overview

This data is on traits of Labrador retrievers based on the owners knowledge of the dog. The data was collected to see how the genetics of the dogs effected their personality. The data is from a questionnaire that ask 102* questions about the dogs and gives a value between 1 and 5. 


Ilska, Joanna et al. (2018), Data from: Genetic characterization of dog personality traits, Dryad, Dataset, https://doi.org/10.5061/dryad.171q5

*The questionnaire is from another study that said it has 102 questions but the main paper mentions that there was a supplemental questionnaire.

Lofgren S E, Wiener P, Blott S C, Sánchez Molano E, Woolliams J A et al. , 2014 Management and personality in the labrador retriever dogs. Appl. Anim. Behav. Sci. 156: 44–53, https://www.sciencedirect.com/science/article/pii/S0168159114001099

the data was relatively clean with the main problem being long names for some variables that I have yet to use. Some of the column names are not clear on what they represent such as Job for what is the dogs job (gun dog, pet, or showdog). Also, there are many NA's in the data set which can mainly be ignored with NA.omit. Another minor issue is that all the data is numerical even when it should be categorical like coat color is represented by 0, 1, 2 instead of black, yellow, chocolate. I just left the data as is.


## Analysis 

```{r, include=FALSE}
rm(list = ls())
library(tidyverse)
library(here)
library(ggfortify)
```

```{r, include=FALSE}
DOG <- read.csv(here("Data", "Secor-Cap-Dap-data.csv"))
names(DOG)[4] ="Job"
```

#### Summary of traits
```{r, echo=FALSE}
summary(DOG)
```

#### Hypothosis #1
The job that a dog is trained for effects how excitable the dog is.


I chose job to look at because it seemed to be the category most effected by human interaction. There are three jobs a dog could have on the survey, 0 = gundog (n = 840), 1 = pet (n = 817), and 2 = showdog (n = 140).

```{r, echo=FALSE, message=FALSE}
ggplot(na.omit(DOG), aes(x = Excite)) + geom_histogram()
```
This figure shows the distribution of the Excite column. The distribution looks likes its normal

```{r, echo=FALSE, message=FALSE}
ggplot(data = na.omit(DOG), aes(x = Excite)) + geom_histogram() + facet_wrap(~Job)
```
This figure shows the distribution of Excite over the three job types. All three distributions look normal but gundogs may have a slight skew to the right. 

```{r, echo=FALSE, message=FALSE}
DOG$Job <- as.factor(DOG$Job)
ggplot(data = na.omit(DOG), aes(group = Job, y = Excite, x = Job)) + geom_boxplot() + geom_point(size = 4, color = 'lightgrey', alpha = 0.5)
```
This figure shows that all three jobs seemingly have a similar level of excitability but showdogs have a lower mean than pets and gundog mean is even lower. 

There will likely be a difference in excitability between the three jobs as there is a noticeable difference between pets and gundogs but whether or not it is significant is yet to be seen.

```{r, echo=FALSE}
jvelm <- lm(Excite ~ Job, data = na.omit(DOG))
autoplot(jvelm)
```
The residual fitted looks very good as the blue line matches the grey dotted line and there are few outlines. Even though it looks a little funny, job is a categorical so the values should be in parallel lines like they are. The points on the normal Q-Q look nearly identical to the grey line except at the top.

```{r, echo=FALSE}
summary(jvelm)
```

There is a significant differences between all three jobs with the biggest difference between gundog and pet and with p = 2e-16. There is a closer but still statistically significant difference between pet and show dog. We fail to reject the hypothesis that a dogs job effects its excitability. Teaching dogs different jobs does seem have an impact on how they act. 


#### Hypothosis #2
A dogs coat color effects its level of excitability.

This is a very similar test to the last one because job was a more nurture variable whereas coat color is a genetic variable.

There are three coat colors, 0 = black (n = 1144), 1 = yellow (n = 521), and 2 = chocolate (n = 310).

We think that it is unlikely there is a significant differences between any of the coat colors.

```{r, echo=FALSE, message=FALSE}
ggplot(data = na.omit(DOG), aes(x = Excite)) + geom_histogram() + facet_wrap(~Coat_colour)
```
This figure shows the ditrobution of Excitement over each coat color and they all seem to be normal.

```{r, echo=FALSE, message=FALSE}
DOG$Coat_colour <- as.factor(DOG$Coat_colour)
ggplot(data = na.omit(DOG), aes(group = Coat_colour, y = Excite, x = Coat_colour)) + geom_boxplot() + geom_point(size = 4, color = 'lightgrey', alpha = 0.5)
```
This figure shows both black and yellow have similar levels of excitability but chocolate labs seemingly are more excitable.

from just the graphs we would expect there to possibly be difference between chocolate and the other two coat colors.

```{r, echo=FALSE}
ccvelm <- lm(Excite ~ Coat_colour, data = na.omit(DOG))
autoplot(jvelm)
```
This figure shows that residual fitted and normal Q-Q look good and have relatively few outliers.

```{r, echo=FALSE}
summary(ccvelm)
```
Between black labs and yellow labs there is not a significant difference in level of excitability p = 0.512 but between black labs and chocolate labs there is a significant differences in level of excitability. We have to fail to reject the hypothesis that coat color does effect excitability in labs. We were not expecting that coat color could change how excitable dogs can be. It would likely be good to check the results of this because chocolate labs were the smallest group to see if there was another variable influencing the data.


#### Hypothosis #3
The health of a dog can effect the excitability in Labrador retrievers.

We wanted to see if there was a third variable that could effect how excitable a dog is and health of a do to see if illness or injury made the dog less excitable. 0 = some health issues (n = 1697) and 1 = no heath issues (n = 278)

We think that there will be no significant change in excitement level mainly because level of injury/illness is not specified so it could be extreme or nothing that impacts the dogs life.

```{r, echo=FALSE, message=FALSE}
DOG$Health <- as.factor(DOG$Health)
ggplot(DOG, aes(group = Health, x = Health, y = Excite)) + geom_boxplot() + geom_point(size = 4, color = 'lightgrey', alpha = 0.5)
```
This graph is a bar and whisker plot comparing dogs with some health issues and dogs with out health issues. The healthy dogs seem to be slightly more excitable as both the mean and Q1 are higher than dogs with health issues.

```{r}
t.test(Excite ~ Health, data = DOG)
```

There is no significant difference between some heath issues and no health issues when it comes to the level of excitement of Labrador retrievers p = 0.2265. We reject the hypothesis. This is no real surprise as health problems vs no health problems is rather vague and because this comes from a survey it is hard to compare scale of injury.


## Take-Home Messages

Too much data can be overwhelming and selecting certain parts and narrowing the focus helps a lot. 
Something I would do different is that I would pick a different data set that was not based on a questionnaire. 
next time that I collect data I would record in a way that each column header is easy to understand and make the categorical data words or words and numbers in a different column so that the data is very easy to understand.


