---
title: "Capstone analysis Bark"
author: "HarrySecor"
date: "11/19/2022"
output: html_document
---

```{r}
rm(list = ls())
library(tidyverse)
library(here)
library(ggfortify)
```

```{r}
DOG <- read.csv(here("Data", "Secor-Cap-Dap-data.csv"))
```

find male vs female by Gender_status %% 2
need to do this for later before Gender_status becomes a factored value

```{r}
DOG$Fixed <- DOG$Gender_status %% 2
```

I used summary to find a characteristic that seems to be relatively evenly distributed with min = 1, mean = 2.5, and max = 5 roughly
```{r}
summary(DOG)
```

histogram where x = Bark

```{r}
ggplot(DOG, aes(x = Bark)) + geom_histogram()
```

histograms where x = Bark and facet_wrap job

```{r}
ggplot(data = na.omit(DOG), aes(x = Bark)) + geom_histogram() + facet_wrap(~WM2)
```

box plot of x = WM2 and y = Bark

```{r}
DOG$WM2 <- as.factor(DOG$WM2)
ggplot(data = na.omit(DOG), aes(y = Bark, x = WM2)) + geom_boxplot() + geom_point(size = 4, color = 'lightgrey', alpha = 0.5)
```

autoplot and summary of linear model with y = Bark

```{r}
jvelm <- lm(Bark ~ WM2, data = na.omit(DOG))
autoplot(jvelm)
summary(jvelm)
```


box plot of x = Coat_colour and y = Bark

```{r}
DOG$Coat_colour <- as.factor(DOG$Coat_colour)
ggplot(data = na.omit(DOG), aes(y = Bark, x = Coat_colour)) + geom_boxplot() + geom_point(size = 4, color = 'lightgrey', alpha = 0.5)
```

autoplot and summary of linear model with x = Coat_colour and y = Bark

```{r}
ccvelm <- lm(Bark ~ Coat_colour, data = na.omit(DOG))
autoplot(ccvelm)
summary(ccvelm)
```


dot plot of x = Age and y = Bark

```{r}
ggplot(DOG, aes(y = Bark, x = Age)) + geom_point() + xlab("Age (Days)")
```

autoplot of linear model with x = Age and y = Bark

```{r}
avelm <- lm(Bark ~ Age, data = DOG)
autoplot(avelm)
```

Anova test and summary of linear model with x = Age and y = Bark

```{r}
anova(avelm)
summary(avelm) # Starts at Call:
```

box plot of x = Health and y = Bark

```{r}
DOG$Health <- as.factor(DOG$Health)
ggplot(DOG, aes(x = Health, y = Bark)) + geom_boxplot() + geom_point(size = 4, color = 'lightgrey', alpha = 0.5)
```

t.test with x = Health and y = Bark

```{r}
t.test(Bark ~ Health, data = DOG)
```

box plot of x = Gender_status and y = Bark

```{r}
DOG$Gender_status <- as.factor(DOG$Gender_status)
ggplot(data = na.omit(DOG), aes(y = Bark, x = Gender_status)) + geom_boxplot() + geom_point(size = 4, color = 'lightgrey', alpha = 0.5)
```

autoplot of linear model with x = Gender_status and y = Bark

```{r}
gsvelm <- lm(Bark ~ Gender_status, data = na.omit(DOG))
autoplot(gsvelm)
summary(gsvelm)
```


box plot of x = Fixed and y = Bark

```{r}
DOG$Fixed <- as.factor(DOG$Fixed)
ggplot(DOG, aes(x = Fixed, y = Bark)) + geom_boxplot() + geom_point(size = 4, color = 'lightgrey', alpha = 0.5)
```

autoplot of linear model with x = Fixed and y = Bark

```{r}
fvelm <- lm(Bark ~ Fixed, data = na.omit(DOG))
autoplot(fvelm)
summary(fvelm)
```

box plot of x = InOut and y = Bark

```{r}
DOG$InOut <- as.factor(DOG$InOut)
ggplot(data = na.omit(DOG), aes(x = InOut, y = Bark)) + geom_boxplot() + geom_point(size = 4, color = 'lightgrey', alpha = 0.5)
```

autoplot of linear model with x = Fixed and y = Bark

```{r}
iovelm <- lm(Bark ~ InOut, data = na.omit(DOG))
autoplot(iovelm)
summary(iovelm)
```

box plot of x = TimeEx and y = Bark

```{r}
DOG$TimeEx <- as.factor(DOG$TimeEx)
ggplot(data = na.omit(DOG), aes(x = TimeEx, y = Bark)) + geom_boxplot() + geom_point(size = 4, color = 'lightgrey', alpha = 0.5)
```


```{r}
tevelm <- lm(Bark ~ TimeEx, data = na.omit(DOG))
autoplot(tevelm)
summary(tevelm)
```


